---
title: "GR5702 - Using SQL with R"
author: "Samir Char, Christodoulos Constantinides"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Setup  
```{r Installing packages, eval=FALSE, echo=TRUE}
install.packages(c("dbplyr", "RMariaDB"))
```
## Connecting  
Replace the database credentials   
```{r installing packages}
library(RMariaDB)
connection <- dbConnect(RMariaDB::MariaDB(), user='root', password='', dbname='lahmansbaseballdb', host='localhost')
``` 
Now connection is responsible for executing queries and other actions related to the database.   
We can view which tables are in the database  
```{r}
dbListTables(connection)
```
## Create - Read - Update - Delete operations
We can also perform CRUD operations  
  
## Insert

```{r Insert}
query <- "INSERT INTO leagues(lgID, league, active) VALUES ('TT', 'Test League', 'N')"
res <- dbSendQuery(connection,query)
res
```  
### Update  
```{r Update}
query <- "UPDATE leagues SET active='Y' WHERE lgID='TT'"
res <- dbSendQuery(connection,query)
res
```
### Delete  
```{r Delete}
query <- "DELETE FROM leagues WHERE lgID='TT'"
res <- dbSendQuery(connection,query)
res
```  
### Select  
```{r Select}
query <- 'SELECT yearID, G_all FROM appearances WHERE playerID="willite01" ORDER BY yearID'
rs <- dbSendQuery(connection,query)
dbRows<-dbFetch(rs)
dbRows
```

## Dataframe cleaning  
Before visualizing our result we should do some cleaning like column type conversion.  
```{r Cleaning the resulting frame}
dbRows$yearID <- as.numeric(dbRows$yearID)
```

## Visualizing our Dataframe  
We are now ready to visualize our Dataframe.
```{r Visualizing our query}
library(tidyverse)
ggplot(dbRows, aes(yearID, G_all)) + 
  geom_line() + ggtitle("Games played by willite01 each year") +
  xlab("Year") + ylab("Games Played")

```

